---
title: "CatAct v3"
author: "Removed for review"
date: "2023-01-17"
output: html_document
---

```{r setup, warning=FALSE, message=F}
library(here)
library(tidyverse)
library(ggplot2)
library(viridisLite)
library(lme4)
library(ggstance)
library(cowplot)
library(AICcmodavg)
library(RColorBrewer)
library(jsonlite)
library(tidyjson)
library(rlang)
library(car)
library(nnet)
library(mlogit)
library(lmerTest)
source("helper.R")

data_path <- here("..","..","data","v3","processed", "catact-v3-alldata-processed.csv")
sampling_data_path <- here("..","..","data","v3","processed", "catact-v3-sampling-data.csv")
test_data_path <- here("..","..","data","v3","processed", "catact-v3-test-data.csv")
test_data_long_path <- here("..","..","data","v3","processed", "catact-v3-test-data-long.csv")
write_path <- here("..","..","data","v3","processed")
figure_path <- here("figures")
```

# Data Processing

## Read in data

```{r, warning=FALSE, message=F}
d <- read_csv(data_path)
sampling_data <- read_csv(sampling_data_path)
test_array_clean <- read_csv(test_data_path)
test_array_options_clean <- read_csv(test_data_long_path)
```

## handle exclusions

```{r, warning=FALSE, message=F}
d <- d %>%
  filter(exclude_participant==0)
sampling_data <- sampling_data  %>%
  filter(exclude_participant==0) %>%
  filter(levenshtein_distance<2)
test_array_clean <- test_array_clean %>%
  filter(exclude_participant==0) %>%
  filter(levenshtein_distance<2)
test_array_options_clean <- test_array_options_clean %>%
  filter(exclude_participant==0) %>%
  filter(levenshtein_distance<2)
```


## Extract category responses
```{r}
category_labels <- d %>%
  filter(!is.na(word_meaning)) %>%
  distinct(subject,current_category_training_level,current_category_label_level,current_category_kind,word_meaning)
write_csv(category_labels,here(write_path,"catact-v3-label-responses.csv"))
  
```

## Summarize Test Data

```{r}
## shorter test representation
subj_test_accuracy <- test_array_clean %>%
  group_by(subject,trial_type,current_category_training_level,current_category_label_level,current_category_kind,final_choice_array,total_number_correct_options) %>%
  summarize(
    total_choices = n(),
    ground_truth_correct_choices = sum(test_choice_type_label_consistent),
    ground_truth_incorrect_choices = sum(test_choice_type_label_consistent==0),
    training_consistent_choices = sum(test_choice_type_training_consistent),
    training_inconsistent_choices = sum(test_choice_type_training_consistent==0),
    training_category_match_subordinate_choices = sum(test_choice_match_category==1 & test_choice_type == "subordinate"),
    training_category_match_basic_choices = sum(test_choice_match_category==1 & test_choice_type == "basic"),
    training_category_match_superordinate_choices = sum(test_choice_match_category==1 & test_choice_type == "superordinate"),
    training_category_match_subordinate_percent = training_category_match_subordinate_choices/2,
    training_category_match_basic_percent = training_category_match_basic_choices/2,
    training_category_match_superordinate_percent = training_category_match_superordinate_choices/4
  )

subj_test_accuracy_long <- subj_test_accuracy %>%
  select(subject,trial_type,current_category_training_level,current_category_label_level,current_category_kind,final_choice_array,total_number_correct_options,training_category_match_subordinate_percent,training_category_match_basic_percent,training_category_match_superordinate_percent) %>%
  group_by(subject,trial_type,current_category_training_level,current_category_label_level,current_category_kind,final_choice_array,total_number_correct_options) %>%
  pivot_longer(cols=training_category_match_subordinate_percent:training_category_match_superordinate_percent,names_to = "choice_type",values_to = "percent_chosen")

#combine with sampling data
subj_test_accuracy_long <- subj_test_accuracy_long %>%
  left_join(sampling_data)

overall_accuracy <- subj_test_accuracy_long %>%
  ungroup() %>%
  group_by(current_category_training_level,choice_type) %>%
  summarize(
    N=n(),
    average_percent=mean(percent_chosen)
  ) %>%
  mutate(
    choice_type = str_remove(choice_type,"training_category_match_"),
    choice_type = str_remove(choice_type,"_percent")
  )

overall_accuracy_label_level <- subj_test_accuracy_long %>%
  ungroup() %>%
  group_by(current_category_training_level,current_category_label_level,choice_type) %>%
  summarize(
    N=n(),
    average_percent=mean(percent_chosen)
  ) %>%
  mutate(
    choice_type = str_remove(choice_type,"training_category_match_"),
    choice_type = str_remove(choice_type,"_percent")
  )

overall_accuracy_label_level_sampling <- subj_test_accuracy_long %>%
  ungroup() %>%
  group_by(sampling_choice_narrowly_constraining,current_category_training_level,current_category_label_level,choice_type,) %>%
  summarize(
    N=n(),
    average_percent=mean(percent_chosen)
  ) %>%
  mutate(
    choice_type = str_remove(choice_type,"training_category_match_"),
    choice_type = str_remove(choice_type,"_percent")
  )

### longer test representation
subj_test_accuracy_all <- test_array_options_clean %>%
  group_by(subject,trial_type,current_category_training_level,current_category_label_level,current_category_kind,final_choice_array,total_number_correct_options_label,total_number_correct_options_training) %>%
  summarize(
    accuracy_training = mean(is_match_to_training, na.rm=TRUE),
    accuracy_label = mean(is_match_to_label,na.rm=TRUE),
    hit_rate_training = mean(hit_training,na.rm=TRUE),
    hit_rate_label = mean(hit_label,na.rm=TRUE),
    false_alarm_rate_training = mean(false_alarm_training,na.rm=TRUE),
    false_alarm_rate_label = mean(false_alarm_label,na.rm=TRUE)
  ) %>%
  mutate(
    hit_rate_training_adj= case_when(
      hit_rate_training==1 ~ 1 - 1/(2*total_number_correct_options_training),
      hit_rate_training==0 ~ 1/(2*total_number_correct_options_training),
      TRUE ~ hit_rate_training
    ),
    hit_rate_label_adj= case_when(
      hit_rate_label==1 ~ 1 - 1/(2*total_number_correct_options_label),
      hit_rate_label==0 ~ 1/(2*total_number_correct_options_label),
      TRUE ~ hit_rate_label
    ),
    false_alarm_rate_training_adj= case_when(
      false_alarm_rate_training==0 ~ 1/(2*total_number_correct_options_training),
      false_alarm_rate_training==1 ~ 1 - 1/(2*total_number_correct_options_training),
      TRUE ~ false_alarm_rate_training
    ),
    false_alarm_rate_label_adj= case_when(
      false_alarm_rate_label==0 ~ 1/(2*total_number_correct_options_label),
      false_alarm_rate_label==1 ~ 1 - 1/(2*total_number_correct_options_label),
      TRUE ~ false_alarm_rate_label
    )
  ) %>%
  mutate(
    dprime_training=qnorm(hit_rate_training_adj) - qnorm(false_alarm_rate_training_adj),
    dprime_label=qnorm(hit_rate_label_adj) - qnorm(false_alarm_rate_label_adj),
    c_training=-.5*(qnorm(hit_rate_training_adj) + qnorm(false_alarm_rate_training_adj)),
    c_label=-.5*(qnorm(hit_rate_label_adj) + qnorm(false_alarm_rate_label_adj)))

subj_test_accuracy_all <- subj_test_accuracy_all %>%
  left_join(sampling_data)

overall_accuracy_label_level_sampling_all <- subj_test_accuracy_all %>%
  ungroup() %>%
  group_by(sampling_choice_narrowly_constraining,current_category_training_level,current_category_label_level) %>%
  summarize(
    N=n(),
    average_dprime_training=mean(dprime_training),
    average_dprime_label=mean(dprime_label)
  ) 
```

# Sampling Phase
 
## Learners flexibly shift their sampling choices depending on the training condition. 

```{r}
sampling_data <- sampling_data %>%
  mutate(
    sampled_category_level_kind_info_choice_order = case_when(
      sampled_category_level_kind_info == "outside_category" ~ "a_outside_category",
      sampled_category_level_kind_info == "sub" ~ "b_sub",
      sampled_category_level_kind_info == "bas" ~ "c_bas",
      sampled_category_level_kind_info == "sup" ~ "d_sup"
  ),
  current_category_training_level_order = case_when(
      current_category_training_level == "narrow" ~ "a_narrow",
      current_category_training_level == "intermediate" ~ "b_intermediate",
      current_category_training_level == "broad" ~ "c_broad"
  )
  )

#fit multinomial model
nnetFixedModelStr <- 'sampled_category_level_kind_info_choice_order ~ current_category_training_level_order'
nnetFixedFit <-  multinom(as.formula(nnetFixedModelStr), sampling_data, maxit=10000, abstol=1e-8, reltol=1e-10)
summary(nnetFixedFit)
Anova(nnetFixedFit,type="III")

# Alternative (equivalent) approach with mlogit
# reformat data for mlogit
mlogit_sampling_data <-  mlogit.data(data=select(sampling_data,subject,sampled_category_level_kind_info_choice_order,current_category_training_level_order,current_category_kind), choice='sampled_category_level_kind_info_choice_order', shape="wide", id.var='subject')

# specify model formula
mlogit_sampling_formula_baseline <- ' sampled_category_level_kind_info_choice_order ~ 0 | 1 | 0'
mlogit_sampling_formula_training <- 'sampled_category_level_kind_info_choice_order ~ 0 | current_category_training_level_order | 0'

# fit models
mlogit_sampling_baseline <- mlogit(as.formula(mlogit_sampling_formula_baseline), data=mlogit_sampling_data, panel=FALSE, iterlim=10000, tol = 1e-8)
print(summary(mlogit_sampling_baseline))
mlogit_sampling_training <- mlogit(as.formula(mlogit_sampling_formula_training), data=mlogit_sampling_data, panel=FALSE, iterlim=10000, tol = 1e-8)
print(summary(mlogit_sampling_training))

# overall test of condition (identical result to nnet method above)
lrtest(mlogit_sampling_baseline,mlogit_sampling_training)
```

### Robustness Checks

#### Controlling for Category Type

```{r}
nnetFixedModelStr <- 'sampled_category_level_kind_info_choice_order ~ current_category_training_level_order+current_category_kind'
nnetFixedFit <-  multinom(as.formula(nnetFixedModelStr), sampling_data, maxit=10000, abstol=1e-8, reltol=1e-10)
summary(nnetFixedFit)
Anova(nnetFixedFit,type="II")
```

#### Adding by-participant random effects

```{r}
#set up random effects and model fitting parameters
parNames=names(mlogit_sampling_training$coefficients)[1:3]
print(parNames)
rparArg=rep('n',length(parNames))
names(rparArg)=parNames
print(rparArg)
NRCRDRAWS=1000

#fit models
if (NRCRDRAWS>0){
mlogit_sampling_baseline_re=mlogit(as.formula(mlogit_sampling_baseline), data=mlogit_sampling_data, rpar=rparArg, 
                panel=TRUE, R=NRCRDRAWS,print.level=1)
print(summary(mlogit_sampling_baseline_re))
}
if (NRCRDRAWS>0){
mlogit_sampling_training_re=mlogit(as.formula(mlogit_sampling_formula_training), data=mlogit_sampling_data, rpar=rparArg, 
                panel=TRUE, R=NRCRDRAWS,print.level=1)
print(summary(mlogit_sampling_training_re))
}

# LR Test
lrtest(mlogit_sampling_baseline_re,mlogit_sampling_training_re)
```

## Participants' sampling choices. 

### Confirming Sampling Choices

```{r}
#consistent
m <- glmer(sampling_choice_consistent_b~offset(logit(chance_consistent))+(1|subject), data=sampling_data, family=binomial,glmerControl(optimizer="bobyqa"))
summary(m)
confint(m,method="Wald")

#effect of training
m <- glmer(sampling_choice_consistent_b~offset(logit(chance_consistent))+current_category_training_level+(1|subject), data=sampling_data, family=binomial,glmerControl(optimizer="bobyqa"))
summary(m)
Anova(m,type="III")
```

### Constraining Sampling Choices

```{r}
#constraining
m <- glmer(sampling_choice_narrow_constraining_b~offset(logit(chance_narrowly_constraining))+(1|subject), data=sampling_data, family=binomial)
summary(m)

m <- glmer(sampling_choice_narrow_constraining_b~offset(logit(chance_narrowly_constraining))+current_category_training_level+(1|subject), data=sampling_data, family=binomial)
summary(m)
Anova(m,type="III")

## Condition-specific follow-ups

sampling_data$current_category_training_level_narrow <- factor(sampling_data$current_category_training_level,levels=c("narrow","intermediate","broad"))
#intercept centered on narrow condition (reference level)
m <- glmer(sampling_choice_narrow_constraining_b~offset(logit(chance_narrowly_constraining))+current_category_training_level_narrow+(1|subject), data=sampling_data, family=binomial)
summary(m)

sampling_data$current_category_training_level_intermediate <- factor(sampling_data$current_category_training_level,levels=c("intermediate","narrow","broad"))
#intercept centered on intermediate condition (reference level)
m <- glmer(sampling_choice_narrow_constraining_b~offset(logit(chance_narrowly_constraining))+current_category_training_level_intermediate+(1|subject), data=filter(sampling_data), family=binomial)
summary(m)

sampling_data$current_category_training_level_broad <- factor(sampling_data$current_category_training_level,levels=c("broad","narrow","intermediate"))
#intercept centered on broad condition (reference level)
m <- glmer(sampling_choice_narrow_constraining_b~offset(logit(chance_narrowly_constraining))+current_category_training_level_broad+(1|subject), data=filter(sampling_data), family=binomial)
summary(m)
```

```{r}
subj_sampling <- sampling_data %>%
  group_by(subject) %>%
  summarize(
    mean_consistent=mean(sampling_choice_consistent_b),
    mean_consistent_adj=mean(sampling_choice_consistent_b-chance_consistent),
    mean_narrowly_constraining = mean(sampling_choice_narrow_constraining_b),
    mean_narrowly_constraining_adj = mean(sampling_choice_narrow_constraining_b-chance_narrowly_constraining),
    mean_outside_category=mean(sampled_category_level_kind_info=="outside_category"))

t.test(subj_sampling$mean_consistent)
t.test(subj_sampling$mean_narrowly_constraining)
t.test(subj_sampling$mean_consistent,subj_sampling$mean_narrowly_constraining,paired=T)
t.test(subj_sampling$mean_consistent_adj,subj_sampling$mean_narrowly_constraining_adj,paired=T)
```

# Test

## Trial-by-trial model

```{r}
m <- glmer(is_chosen ~ current_category_training_level + test_image_match_category + (1+current_category_training_level+ test_image_match_category|subject)+(1|test_image), data=test_array_options_clean,family=binomial,glmerControl(optimizer="bobyqa"))
summary(m)
Anova(m,type="III")
```

## Training condition modulated participants’ choices at test (Proportion choices)

```{r}
subj_test_prop <- test_array_options_clean %>%
  group_by(subject,current_category_training_level,current_category_label_level,current_category_kind) %>%
  summarize(
    N_subordinate = sum(test_image_type=="subordinate" & test_image_match_category==1),
    prop_subordinate = sum(test_image_type=="subordinate" & test_image_match_category==1 & is_chosen == 1)/sum(test_image_type=="subordinate" & test_image_match_category==1),
    N_basic = sum(test_image_type=="basic" & test_image_match_category==1),
    prop_basic = sum(test_image_type=="basic" & test_image_match_category==1 & is_chosen == 1)/sum(test_image_type=="basic" & test_image_match_category==1),
    N_superordinate = sum(test_image_type=="superordinate" & test_image_match_category==1),
    prop_superordinate = sum(test_image_type=="superordinate" & test_image_match_category==1 & is_chosen == 1)/sum(test_image_type=="superordinate" & test_image_match_category==1)
  )

overall_test_prop <- subj_test_prop %>%
  group_by(current_category_training_level) %>%
  summarize(
    N=n(),
    subordinate_percent = mean(prop_subordinate),
    basic_percent = mean(prop_basic),
    superordinate_percent = mean(prop_superordinate)
  )

# fit individual lmer models
## subordinate choices
m_sub <- lmer(prop_subordinate~current_category_training_level+(1|subject),data=subj_test_prop)
summary(m_sub)
Anova(m_sub,type="III")

## basic choices
m_bas <- lmer(prop_basic~current_category_training_level+(1|subject),data=subj_test_prop)
summary(m_bas)
Anova(m_bas,type="III")
### ref level as Narrow condition
subj_test_prop$current_category_training_level_narrow <- factor(subj_test_prop$current_category_training_level, levels=c("narrow","intermediate","broad"))
m_bas_narrow <- lmer(prop_basic~current_category_training_level_narrow+(1|subject),data=subj_test_prop)
summary(m_bas_narrow)

## superordinate choices
m_sup <- lmer(prop_superordinate~current_category_training_level+(1|subject),data=subj_test_prop)
summary(m_sup)
Anova(m_sup,type="III")

```

# Relationship between Sampling and Test

## Consistent choices and test performance

```{r}
subj_test_accuracy_all <- subj_test_accuracy_all %>%
  left_join(subj_sampling)

ggplot(subj_test_accuracy_all,aes(mean_consistent,dprime_label))+
  geom_point()+
  geom_smooth(method="lm",se=FALSE,aes(color=current_category_training_level))+
  geom_smooth(method="lm")

# overall effect
m <- lmer(dprime_label~ mean_consistent+(1+mean_consistent|subject),data=subj_test_accuracy_all)
summary(m)

# interaction with Training Condition
m <- lmer(dprime_label~ mean_consistent*current_category_training_level+(1+mean_consistent|subject),data=subj_test_accuracy_all)
summary(m)
Anova(m,type="II")
```

## Constraining choices and test performance

```{r}
ggplot(subj_test_accuracy_all,aes(mean_narrowly_constraining,dprime_label))+
  geom_point()+
  geom_smooth(method="lm",se=FALSE,aes(color=current_category_training_level))+
  geom_smooth(method="lm")

#Overall effect
m <- lmer(dprime_label~ mean_narrowly_constraining+(1+mean_narrowly_constraining|subject),data=subj_test_accuracy_all)
summary(m)

#interaction with Training Condition
m <- lmer(dprime_label~ mean_narrowly_constraining*current_category_training_level+(1+mean_narrowly_constraining|subject),data=subj_test_accuracy_all)
summary(m)
Anova(m,type="II")
```

